Measuring contrast processing in the visual system using SSVEP

### 2024-04-20 review.qmd
Fixed equation error
A few minor tweaks

# Now fit a linear regression to the first 4 contrast levels (before it saturates) and find out where this crosses the Y-axis
X = np.log10(contrastRange[:4]).reshape(-1, 1)
y = transducedSignal[:4]
reg = LinearRegression().fit(X, y)
x_intercept = 10 ** (-reg.intercept_ / reg.coef_[0])

axs.set_title(f'Simulated sweep VEP at {nLevels} different contrasts', fontsize=10)
axs.semilogx(contrastRange, transducedSignal, 'o')
axs.semilogx(contrastRange[:4], reg.predict(X), 'r--', label='Linear Fit')
axs.semilogx([x_intercept], [0], 'ro', label='Threshold')
axs.set_xlabel('Contrast')
axs.set_ylabel('Amplitude (a.u.)')
axs.set_xlim(0.01, 1)
axs.set_ylim(0, 10)
axs.legend()
axs.grid()
plt.tight_layout()
plt.savefig('Figures/sweepplot.pdf')

@@

plt.subplot(2, 2, 3)  # Assuming this is the third plot for Autism
plt.plot(finecontrastsdB, respASD, lw=4, color='darkgreen', label='ASD children')
plt.plot(finecontrastsdB, respCON, lw=4, color='black', linestyle='--', label='Control children')
plt.xticks(ticklocsx, ticklabelsx)
plt.yticks(np.arange(0, 1.6, 0.5))
plt.xlabel('Target contrast (%)')
plt.ylabel('SSVEP response')
plt.title('(c) Autism')
plt.legend()
plt.grid()

@@

plt.subplot(2, 2, 4)  # Fourth plot for Parkinson's
plt.plot(finecontrastsdB, respParkPatients, lw=4, color='red', label='hLRRK2 transgenic flies')
plt.plot(finecontrastsdB, respParkControls, lw=4, color='black', linestyle='--', label='TH+ (control) flies')
plt.xticks(ticklocsx, ticklabelsx)
plt.yticks(np.arange(0, 7, 2))
plt.xlabel('Target contrast (%)')
plt.ylabel('SSVEP response')
plt.title("(d) Parkinson's")
plt.legend()
plt.grid()

@@

where $C_\mathrm{mask}}$ reflects the contrast of the mask component at a distinct frequency from that of the target. The effect of this extra term is to reduce the target response (see right panel of @fig-twoinputs). For a linear contrast axis, the contrast response function becomes shallower, whereas on a logarithmic contrast axis it maintains its steepness and shifts to the right.

@@

![Sweep VEP simulation showing how a contrast detection threshold can be estimated from sweep VEP data measured at higher contrasts.](Figures/sweepplot.pdf){#fig-sweepvep}

@@

Amblyopia is a disorder of binocular vision, characterised by one eye contributing much less to perception than the other. This is often due to strabismus (squint) or anisometropia (difference in optical prescription between the eyes) during development. Contemporary accounts suggest that the amblyopic eye is suppressed by signals from the fellow eye. SSVEPs provide a convenient and objective method to characterise the difference in neural response to signals in each eye, and typically show reduced responses to stimuli in the amblyopic eye across the contrast range [@Baker2015; @Lygo2021]. There are currently many novel binocular treatments for amblyopia under development, often involving virtual reality or stereo display systems designed to encourage the two eyes to work together. The steady-state approach may be more sensitive and objective than typical acuity measurements, and also has the potential to measure suppression between the eyes directly [e.g. @Zheng2019; @Du2023; @Hu2023].

>>>>

# Now fit a linear regression to the first 4 contrast levels (before it saturates) and find out where this crosses the Y-axis
X = np.log10(contrastRange[:4]).reshape(-1, 1)
X2 = np.log10(np.logspace(0.1,1.1,nLevels)/100).reshape(-1, 1)
y = transducedSignal[:4]
reg = LinearRegression().fit(X, y)
x_intercept = 10 ** (-reg.intercept_ / reg.coef_[0])

axs.set_title(f'Simulated sweep VEP at {nLevels} different contrasts', fontsize=10)
axs.semilogx(contrastRange, transducedSignal, 'o')
axs.semilogx(contrastRange[:4], reg.predict(X), 'r-', label='Linear Fit')
axs.semilogx(np.logspace(0.1,1.1,nLevels)/100, reg.predict(X2), 'r--')
axs.semilogx([x_intercept], [0], 'ro', label='Threshold')
axs.set_xlabel('Contrast')
axs.set_ylabel('Amplitude (a.u.)')
axs.set_xlim(0.01, 1)
axs.set_ylim(0, 10)
axs.legend()
axs.grid()
plt.tight_layout()
plt.savefig('Figures/sweepplot.pdf')

@@

plt.subplot(2, 2, 3)  # Assuming this is the third plot for Autism
plt.plot(finecontrastsdB, respASD, lw=4, color='darkgreen', label='ASD children')
plt.plot(finecontrastsdB, respCON, lw=4, color='black', linestyle='--', label='Control children')
plt.xticks(ticklocsx, ticklabelsx)
plt.yticks(np.arange(0, 2, 0.5))
plt.xlabel('Target contrast (%)')
plt.ylabel('SSVEP response')
plt.title('(c) Autism')
plt.legend()
plt.grid()

@@

plt.subplot(2, 2, 4)  # Fourth plot for Parkinson's
plt.plot(finecontrastsdB, respParkPatients, lw=4, color='red', label='hLRRK2 transgenic flies')
plt.plot(finecontrastsdB, respParkControls, lw=4, color='black', linestyle='--', label='TH+ (control) flies')
plt.xticks(ticklocsx, ticklabelsx)
plt.yticks(np.arange(0, 20, 4))
plt.xlabel('Target contrast (%)')
plt.ylabel('SSVEP response')
plt.title("(d) Parkinson's")
plt.legend()
plt.grid()

@@

where $C_\mathrm{mask}$ reflects the contrast of the mask component at a distinct frequency from that of the target. The effect of this extra term is to reduce the target response (see right panel of @fig-twoinputs). For a linear contrast axis, the contrast response function becomes shallower, whereas on a logarithmic contrast axis it maintains its steepness and shifts to the right.

@@

![Sweep VEP simulation showing how a contrast detection threshold can be estimated from sweep VEP data measured at higher contrasts. The solid line is the regression fit to the lowest four data points, and the dashed line extrapolates the fit back to determine the contrast value when y=0, which gives a threshold estimate (red point).](Figures/sweepplot.pdf){#fig-sweepvep}

@@

Amblyopia is a disorder of binocular vision, characterised by one eye contributing much less to perception than the other. This is often due to strabismus (squint) or anisometropia (difference in optical prescription between the eyes) during development. Contemporary accounts suggest that the amblyopic eye is suppressed by signals from the fellow eye. SSVEPs provide a convenient and objective method to characterise the difference in neural response to signals in each eye, and typically show reduced responses to stimuli in the amblyopic eye (see [@fig-plotclinical]b) across the contrast range [@Baker2015; @Lygo2021]. There are currently many novel binocular treatments for amblyopia under development, often involving virtual reality or stereo display systems designed to encourage the two eyes to work together. The steady-state approach may be more sensitive and objective than typical acuity measurements, and also has the potential to measure suppression between the eyes directly [e.g. @Zheng2019; @Du2023; @Hu2023].
